\ripple: Differentiable Waveforms for Gravitational Wave Data Analysis

### 2023-01-04 src/tex/ms.tex
Edited/added plots and added some new text

A primary issue with both MCMC and nested sampling is that neither utlizes information about the likelihood's derivative and must therefore randomly walk towards areas of highest likelihood.
Derivative based samplers, on the other hand, have been shown to extrapolate well to higher dimensions although they sometimes come with their own drawbacks.
Here we simply aim to demonstrate the utility of a derivative based sampler and their efficiency on a small test problem.
In particular, we will show that proposals made using an Hamiltonian Monte Carlo (MC) sampler are accepted at a much higher rate than a traditional MCMC algorithm.

For our basic example we perform an injection recovery test on a seven dimensional parameter space with the same three detector network setup considered in \S~\ref{subsec:fisher}.
More specifically, we generate Gaussian noise consistent with the measured PSDs for each detector and then inject a BBH signal with parameters: $M_c = 23.82\,\mathrm{M_\odot}$,  $\eta = 0.248$, $\chi_1=0.3$, $\chi_1=-0.4$, $D = 1.6\,\mathrm{Gpc}$, $t_c = 0.0$, and $\phi_c = 0.0$.
Using a standard Gaussian likelihood, we then run the HMC sampler implemented in \flowMC for XX steps. 
\te{Describe more carefully what we do}

@@

>>>>

A primary issue with both MCMC and nested sampling is that neither utlizes information about the likelihood's derivative and must therefore randomly walk towards areas of highest likelihood.
Derivative based samplers, on the other hand, have been shown to extrapolate well to higher dimensions although they sometimes come with their own drawbacks.
Here we simply aim to demonstrate the utility of a derivative based sampler and their efficiency on a small test problem.
% In particular, we will show that proposals made using an Hamiltonian Monte Carlo (MC) sampler are accepted at a much higher rate than a traditional MCMC algorithm.
In particular, we will show that the autocorrelation of a Hamiltonian Monte Carlo (HMC) sampler is significantly lower than a traditional MCMC algorithm.

For our basic example we perform an injection recovery test on a seven dimensional parameter space with the same three detector network setup considered in \S~\ref{subsec:fisher}.
More specifically, we generate Gaussian noise consistent with the measured PSDs for each detector and then inject a BBH signal with parameters: $M_c = 23.82\,\mathrm{M_\odot}$,  $\eta = 0.248$, $\chi_1=0.3$, $\chi_1=-0.4$, $D = 1.6\,\mathrm{Gpc}$, $t_c = 0.0$, and $\phi_c = 0.0$.\footnote{
    The remaining parameters are set to $\pi/3$.
}
Using a standard Gaussian likelihood, we then run the HMC sampler implemented in \flowMC for 200000 steps and the Gaussian random walk sampler for 1500000 steps (each with four randomly initialized independent chains). 
The number of steps and and mass matrix used for each example was hand tuned to give good performance for the specific sampler.
\te{Describe more carefully what we do}

@@

\begin{figure}[t]
	\script{autocorrelation.py}
	\begin{centering}
		\includegraphics[width=\linewidth]{figures/autocorrelation.pdf}
		\caption{
			Autocorrelation
		}
		\label{fig:autocorrelation}
	\end{centering}
\end{figure}
